---
title: "R for Data Science"
author: "Justin Minsk"
date: "September 1, 2017"
output: html_document
---

## Purpose

This site is to play with the Rmarkdown platform more using the code and data from R for Data Science by Hadley Wickman and Garrett Grolemund.

##Data Visulzation

The package I am using for ggplots2 and dyplr.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```
###Chapter one
First set of data.
```{r}
mpg
```
Some different graphs using ggplots.
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) 

#Basic point plot with color shows the engine type and highway milage the car gets and also what type of car it is with color

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth() 

#adds a line of best fit showing that highway mileage lowers with the amount of cylinders 

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)

#breaks data into different graphs based on the class of car

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth(
    data = filter(mpg, class == "subcompact"),
    se = FALSE
  )

#the smooth line now only is based off of the subcompact car class
```
###Chapter Two
Second set of data
```{r}
diamonds
```
Examples of graphs using ggplots
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut)) 
#creates a basic bar graph

ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
#makes the same bar graph different code

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
#makes a bar graph of proportions rather than bulk numbers

ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
#makes a box graph here comparing cut and depth they are found at

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut, fill = cut))
#color adds an outline, fill fills thebars a color

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
#if you use a second varible will fill the bars in proportion to the bar, here cut broken down into clarity

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity),
  position = "fill"
  )
#makes the bars even but shows proportion of cut compared to clarity better

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity),
           position = "dodge"
  )
#makes seprate bars to better show numbers over proportions

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), postion = "jitter")
#makes it so each point is represented, but with loss of accuracy


ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot() +
  coord_flip()
#makes a box plot of class compared to highway milage, coord_flip() flips the axis
```


##Data Transformation
Note that teh veiw() command will not open a window in the browser, but in Rstudio it will open a seprate window to veiw all of the data. 
###Chapter Three
Use this package for the dataset that will be used
```{r message=FALSE, warning=FALSE}
library(nycflights13)
```
The data set shown in table format two different ways
```{r}
nycflights13::flights
#our data set

View(flights)
#opens a window so you can see all the data not just an example 
```
Some examples of the filter command
```{r}
filter(flights, month == 1, day == 1)
#filters the data by first month first day only

(jan1 <- filter(flights, month == 1, day == 1))
#saves that filter into varible jan1

nov_dec <- filter(flights, month %in% c(11, 12))
#assigns all of the nov and dec flights into a table

View(nov_dec)
#test to see that it does
```
Some examples of arrange command
```{r}
View(arrange(flights, year, month, day))
#arranges from low to high by year, month, and day

arrange(flights, desc(arr_delay))
#aranges from high to low the delay, note na or no vaule data points are moved to the end
```
Some examples of select command
```{r}
select(flights, year, month, day)
#selects only certain columns, can use for example (year:day) all columns between two columns or -(year:day) to
#selects everything but the columns you specifiy in this example day, month, and year
#select also uses common string functions like contains, starts_with, etc.

flights_sml <- select(flights, year:day, ends_with("delay"), distance, air_time)
#makes a sample set of day, month, year, all delay sets, distance, and air time

```
Some examples of mutate command
```{r}
mutate(flights_sml, gain = arr_delay - dep_delay, speed = distance / air_time * 60)
#creates columns for gain and speed based off of our parameters

transmute(flights, gain = arr_delay - dep_delay, hours = air_time / 60, gain_per_hour = gain/hours )
#only keeps new varibles that you create using your parameters

```
Some basic commands
```{r}
x <- 1:10
#make x 1-10
lag(x)
#looks at x from one spot the the left so NA to 9 instead
lead(x)
#looks at x from one spot the the right so 2 to NA instead
cumsum(x)
#adds all values
cummean(x)
#finds rolling mean
min_rank(x)
#ranks from smallest to largest 

```
Some examples of using these commands in tandum 
```{r message=FALSE, warning=FALSE}
by_day <- group_by(flights, year, month, day)
#group by_day in flights by year, month, day
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
#finds the mean delay for a day exculding NA's

by_dest <- group_by(flights, dest)
delay <- summarize(by_dest, count = n(), dist = mean(distance, na.rm = TRUE), delay = mean(arr_delay, na.rm = TRUE))
delay <- filter(delay, count > 20, dest != "HNL")
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
#groups flights by destionation, summarizes to coute distance, average delay, and # of flights
#filter to remove short flights and honolulu since it is an outliear

delays <- flights %>%
  group_by(dest) %>%
  summarize(count = n(), dist = mean(distance, na.ra = TRUE), delay = mean(arr_delay, na.ra = TRUE)) %>%
  filter(count > 20, dest != "HNL")
#%>% means or can be thought of as then so create a varible based on our data THEN group the data by destionation
#THEN create a summary with a count and dist as a mean of the distances and delay as the mean of delays from short to highest
#THEN filter out our short flights and Honolulu since it is an outliear

```

